---
title: "Schedule"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

Course material will be uploaded here throughout the semester!

# Rough Schedule

* Course introduction and Parsimony 28.04
* Likelihood and substitution models 05.05
* No Class (Paul Lewis's seminars) 12.05
* Bayesian tree inference 19.05 
* Clock models 31.06
* Tree models 02.06
* Model selection 09.06
* Diversification rate estimation with Joëlle Barido-Sottani 14.06
* Introduction to projects 21.06

# Week 1 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class01.pdf","exercise-01.html")

dt <- data.frame(schedule = c("Slides", "Exercise"),
                 topics = c("tree thinking, parsimony, statistical consistency", "tree building using parsimony"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

# Week 2 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class02.pdf","exercise-02.html","exercise-03.html")

dt <- data.frame(schedule = c("Slides", "Exercise 1", "Exercise 2"),
                 topics = c("maximum likelihood, substitution models", 
                            "tree building using maximum likelihood",
                            "Introduction to RevBayes and the Rev language"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

#Further reading:
#* RevBayes
#* Graphical modelling

```

Don't forget next weeks exercises!!

# Week 3 materials

**No class**

This week Laura and Rachel are off on adventure so there will be no class. During this time we'd like you to watch the following 4 video lectures by [**Paul Lewis**](https://plewis.github.io/) and answer a set of questions we have prepared. Paul Lewis has made numerous important contributions to the field of statistical phylogenetics, as well as producing excellent tools and materials for learning about phylogenetic theory.
The videos we'll use were developed as part of the [phyloseminar](http://phyloseminar.org) series. 

The videos will take about 3.5 hours to watch. Read the questions before you watch each video and write down your answers -- we'll go over these in week 4. Remember **the goal is not to understand everything**! 
Also make note of any questions you have after watching these and we'll try to answer them.

**Questions will be uploaded 06.05.22**

```{r, echo = FALSE}
library(kableExtra)

urls <- c("https://www.youtube.com/watch?v=1r4z0YJq580", "https://www.youtube.com/watch?v=UsLeY0wZr4Y", "https://www.youtube.com/watch?v=4PWlnNsfz90&t=45s", "https://www.youtube.com/watch?v=TLtOS--YwkU",
          "questions-01.html", "questions-02.html", "questions-03.html","questions-04.html")

dt <- data.frame(videos = c("Part 1", "Part 2", "Part 3a", "Part 3b"),
                 topics = c("trees and likelihood", "trees, likelihood and rate heterogeneity", "Intro Bayesian phylogenetics", "Intro Bayesian phylogenetics"),
                 questions = c("Questions 1", "Questions 2", "Questions 3", "Questions 4"))

dt$videos[1] <- paste0("[", dt$videos[1], "](", urls[1], ")")
dt$questions[1] <- paste0("[", dt$questions[1], "](", urls[5], ")")
dt$videos[2] <- paste0("[", dt$videos[2], "](", urls[2], ")")
dt$questions[2] <- paste0("[", dt$questions[2], "](", urls[6], ")")
dt$videos[3] <- paste0("[", dt$videos[3], "](", urls[3], ")")
dt$questions[3] <- paste0("[", dt$questions[3], "](", urls[7], ")")
dt$videos[4] <- paste0("[", dt$videos[4], "](", urls[4], ")")
dt$questions[4] <- paste0("[", dt$questions[4], "](", urls[8], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

#Further reading:

#* [Learning statistics with R: Chapter 1 and Chapter 11](https://learningstatisticswithr.com/book/why-do-we-learn-statistics.html)
#* [tinystats.github.io/teacups-giraffes-and-statistics](https://tinystats.github.io/teacups-giraffes-and-statistics/)
```

# Week 4 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class04.pdf","exercise-04.html")

dt <- data.frame(schedule = c("Slides", "Exercise 4"),
                 topics = c("Follow up on Paul Lewis lectures", 
                            "Bayesian tree inference"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


# Week 5 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class05.pdf","exercise-05.html")

dt <- data.frame(schedule = c("Slides", "Exercise 5"),
                 topics = c("Bayesian timetree inference", 
                            "Time tree inference in RevBayes"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


# Week 6 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class06.pdf","exercise-06.html")

dt <- data.frame(schedule = c("Slides", "Exercise 6"),
                 topics = c("The fossilised birth-death process", 
                            "Dating under the FBD model in RevBayes"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

# Week 7 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class07.pdf","exercise-07.html")

dt <- data.frame(schedule = c("Slides", "Exercise 7"),
                 topics = c("Morphological evolution, model testing and model adequacy", 
                            "Model adequacy in RevBayes"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```



# Week 8 materials

Today we are joined by special celebrity guest and diversification rate model expert [**Joëlle Barido-Sottani**](https://bjoelle.github.io/).

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class08-JBS.pdf","https://revbayes.github.io/tutorials/sse/bisse.html")

dt <- data.frame(schedule = c("Slides", "Exercise 8"),
                 topics = c("Multi-type birth-death processes", 
                            "SSE models in RevBayes"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```



# Week 9 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/class09.pdf","projects.html")

dt <- data.frame(schedule = c("Slides", "Projects page"),
                 topics = c("Phylogenetic comparative methods", 
                            "Projects"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

