---
title: "Schedule"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

Course material will be updated here throughout the semester!

# Rough Schedule

1. Course introduction and parsimony 16.04
2. Likelihood and substitution models 23.04
3. Bayesian inference and MCMC I 30.04
4. Bayesian inference and MCMC II 07.04
5. Clock models 14.05
6. The fossilised birth-death process 21.05
7. Morphological models 28.05
8. Phylogenetic comparative methods 04.06
9. Diversification rates and intro to projects 11.06

# Week 1 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/01-intro.pdf","exercise-01.html")

dt <- data.frame(schedule = c("Slides", "Exercise 1"),
                 topics = c("tree thinking, parsimony", "tree building using parsimony"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Homework**

* Re-familiarise yourself with [how to read a phylogenetic tree](https://artic.network/how-to-read-a-tree.html) and [rooting](https://slides.filogeneti.ca/include/rooting.html)
* Watch [A Brief History of Computational Phylogenetics](https://www.youtube.com/watch?v=PcD15i7yzJ8)

\
\

