---
title: "Schedule"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

Course material will be updated here throughout the semester!

# Rough Schedule

1. Course introduction and parsimony 16.04
2. Likelihood and substitution models 23.04
3. Bayesian inference and MCMC I 30.04
4. Bayesian inference and MCMC II 07.04
5. Bayesian inference and MCMC III 14.05
6. No class! (the Berg) 21.05
7. Molecular dating 28.05
8. Phylogenetic comparative methods 04.06
9. Morphological models and model adequacy 11.06
10. Diversification rates and intro to projects 18.06

# Week 1 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/01-intro.pdf","exercise-01.html")

dt <- data.frame(schedule = c("Slides", "Exercise 1"),
                 topics = c("tree thinking, parsimony", "tree building using parsimony"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Homework**

* Re-familiarise yourself with [how to read a phylogenetic tree](https://artic.network/how-to-read-a-tree.html) and [rooting](https://slides.filogeneti.ca/include/rooting.html)
* Watch [A Brief History of Computational Phylogenetics](https://www.youtube.com/watch?v=PcD15i7yzJ8)


# Week 2 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/02-likelihood.pdf","exercise-03.html")

dt <- data.frame(schedule = c("Slides", "Exercise 2"),
                 topics = c("maximum likelihood, substitution models", 
                            "tree building using maximum likelihood"))
#dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Homework**

* Install the software [RevBayes](https://revbayes.github.io/download)
* Watch [Phyloenetics primer part 1](https://www.youtube.com/watch?v=1r4z0YJq580) by Paul Lewis and answer the [associated questions](questions-01.html)

# Week 3 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/03-Bayesian.pdf","exercise-02.html", "exercise-04.html")

dt <- data.frame(schedule = c("Slides", "Exercise 3", "Exercise 4"),
                 topics = c("Intro to Bayesian inference", 
                            "Intro to RevBayes and the Rev language", "Bayesian tree inference"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Homework**

* Install the software [FigTree](http://tree.bio.ed.ac.uk/software/figtree/) and [Tracer](https://github.com/beast-dev/tracer/releases/tag/v1.7.2)
* Watch [Phyloenetics primer part 2](https://www.youtube.com/watch?v=UsLeY0wZr4Y&t=1614s) by Paul Lewis and answer the [associated questions](questions-02.html)
* Reading [Bayesian Inference of Phylogeny and Its Impact on Evolutionary Biology](https://www.science.org/doi/epdf/10.1126/science.1065889)


# Week 4-5 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/04-Bayesian-II.pdf", "exercise-04.html")

dt <- data.frame(schedule = c("Slides", "Exercise 4"),
                 topics = c("Bayesian inference II", 
                            "Bayesian tree inference"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Homework**

* Watch [Phyloenetics primer part 3a](https://www.youtube.com/watch?v=4PWlnNsfz90) by Paul Lewis and answer the [associated questions](questions-03.html)

**Week 6 - no class!**

# Week 7 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/05-clocks.pdf","exercise-05.html")

dt <- data.frame(schedule = c("Slides", "Exercise 5"),
                 topics = c("Bayesian timetree inference", 
                            "Time tree inference in RevBayes"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Homework**

* Watch [Phyloenetics primer part 3b](https://www.youtube.com/watch?v=TLtOS--YwkU) by Paul Lewis and answer the [associated questions](questions-04.html)

# Week 8 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/Intro_to_PCMs.pdf","https://github.com/emmadunne/PCMs")

dt <- data.frame(schedule = c("Slides", "Exercise 6"),
                 topics = c("Intro to PCMs", 
                            "PCMs using R"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


# Week 9 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/Morphological_models.pdf","https://revbayes.github.io/tutorials/pps_morpho/", "scripts/exercise_morpho.zip")

dt <- data.frame(schedule = c("Slides", "Exercise 7", "Scripts"),
                 topics = c("Morphological models", 
                            "Model adequacy using RevBayes", ""))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```


# Week 10 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/08-FBD.pdf","exercise-07a.html", "exercise-07b.html")

dt <- data.frame(schedule = c("Slides", "Exercise", "Exercise"),
                 topics = c("The fossilised birth-death process", "Exploring the FBD model assumptions", "Dating under the FBD model in RevBayes"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")
dt$schedule[3] <- paste0("[", dt$schedule[3], "](", urls[3], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

\
\

