---
title: "Schedule"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\

Course material will be updated here throughout the semester!

# Rough Schedule

1. Course introduction and Parsimony 16.04
2. Introduction to statistical modelling 23.04
3. Likelihood and substitution models 30.04
4. Bayesian inference and MCMC 07.04
5. Clock models 14.05
6. The fossilised birth-death process 21.05
7. Morphological models 28.05
8. Phylogenetic comparative methods 04.06
9. Introduction to projects 11.06

# Week 1 materials

```{r, echo = FALSE}
library(kableExtra)

urls <- c("slides/week_01.pdf","exercise-01.html")

dt <- data.frame(schedule = c("Slides", "Exercise 1"),
                 topics = c("tree thinking, parsimony", "tree building using parsimony"))
dt$schedule[1] <- paste0("[", dt$schedule[1], "](", urls[1], ")")
dt$schedule[2] <- paste0("[", dt$schedule[2], "](", urls[2], ")")

kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

**Homework**

* Re-familiarise yourself with [how to read a phylogenetic tree](https://artic.network/how-to-read-a-tree.html) and [rooting](https://slides.filogeneti.ca/include/rooting.html)
* Watch [A Brief History of Computational Phylogenetics](https://www.youtube.com/watch?v=PcD15i7yzJ8)

\
\

